---
title: "SO5012 Analysising Data in the Real World  \nMultinomial regression  \nSolutions and commentary"
output: 
  pdf_document: 
    toc: yes
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list = ls())

#whereeveryousavestuff <- "a place"
# Note to RP - Just sneakily saving a variable that has the directory path so make the code look better below
```

## Introduction

This seminar sheet is intended as a introduction to multinomial regression and is a combination of code and **interpretation** for the worksheet *SO5012_semX_multinomial_regression.docx*. 

As with all the previous weeks, first we need to:
1. Set the working directory
2. Load the packages we'll be using
3. Load the data

Here we'll use the `results='hide', message=FALSE` command on the r chunk so that our output is not filled up by this set up code, although the code will be visible. 

```{r loading, results='hide', message=FALSE}

# setwd(whereeveryousavestuff)
# Note to RP - this isnt needed for a project, but you'll need to change if just posting rmd

if (!require(nnet)) install.packages("nnet")
library(nnet)

ukvote2010 <- read.csv("data/ukvote2010.csv")

```

To continue with the preparatory work, and **before** we dive into answering the questions, we first need to look at our data and see what it is, perhaps with some data cleaning if needed.

```{r inspection and cleaning}

str(ukvote2010)

names(ukvote2010)
# we c an that "VoteIntention", "Region", "Gender", "Ethnicity", "MaritalStatus", 
# "Housing", "Qualifications" are character variables - lets set them to factors

# and lets be smart and do it in a loop

factorvars <- c("VoteIntention", "Region", "Gender", "Ethnicity", 
                "MaritalStatus", "Housing", "Qualifications")
for (v in factorvars) {
  ukvote2010[[v]] <- as.factor(ukvote2010[[v]])
  print(table(ukvote2010[[v]], useNA = "ifany"))
}

summary(ukvote2010)
# Now that we have set the factor variables correctly, 
# this short cut will display most of the descriptives, but not the region
# as there are too many levels
table(ukvote2010$Region, useNA = "ifany")

# bt lets pay special attention to the VotingIntention variable, as this 
# will be our depdendent

table(ukvote2010$VoteIntention, useNA = "ifany")
prop.table(table(ukvote2010$VoteIntention, useNA = "ifany"))

```

Here we can see, after setting the factor variables correctly, that none of the questions have any missing values, except for the voting intention variable, where `r round(100*prop.table(table(ukvote2010$VoteIntention, useNA = "ifany"))[[5]],1)`% are missing values - presumably people who did not express an firm opinion (but to confirm this we'd need to look at the data documentation, which you do not have)

\newpage

## Question 1

*Cross-tabulate the variable VoteIntention with the variable Qualifications, setting the table to include the conditional probabilities of VoteIntention, given Qualifications. Which parties do better among those with higher qualification levels and which do worse?*

```{r question 1}

table(ukvote2010$Qualifications,ukvote2010$VoteIntention)

# as this is two way table, we need to tell R which direction to calculate the 
# percentages - this is done by the ,1 at the end of the prop.table command
prop.table(table(ukvote2010$Qualifications,ukvote2010$VoteIntention),1)

# and we can turn it into percentages and round...
round(100*prop.table(table(ukvote2010$Qualifications,ukvote2010$VoteIntention),1),1)

```

Tracing the vote shares (fractions) for each party, we see that Labour’s share is steady at around 25-28%, regardless of education levels. The Conservative share declines with increasing education, from 41% at the lowest levels to 34% at the highest levels.  The Liberal Democratic share increases with higher education, rising from 18 to 30%, while those supporting other parties declines from 13% to 9% as education levels increase.

It is worth noting two things:
1. that these statistics are based on 2010 data, before the coalition government that introduced austerity, and when the Liberal Democrats had their first experience of national government. In short, their vote share was *very* different to what is seen now. 
2. a GCE - a General Certificate of Education Advanced Level - is commonly known as an A-level but this also includes equivalents (BTECs and the like)

\newpage

## Question 2

*Fit a multinomial logistic regression model for VoteIntention, with Labour as the baseline outcome category, using only the variable Qualifications as an explanatory (factor) variable. How can we see from the coefficients which parties do better among those with higher education levels and which do worse? Check that you see the same general patterns as you saw when you cross-tabulated the same data.*

```{r basic mn logit}

levels(ukvote2010$VoteIntention)

# so first we need to relevel Vote Intention 
ukvote2010$VoteIntention <- relevel(ukvote2010$VoteIntention, ref = "Labour")

m1 <- multinom(VoteIntention~ Qualifications, data = ukvote2010)
summary(m1)

```
This stores and displays the coefficients and the *standard errors* (a little like standard deviation, but how much variation there is in the estimate of a coefficient within a model). This needs to be used to calculate the p-values, as the `multinom` package does not do it automatically!

```{r first p values}

zvals <- coef(m1)/summary(m1)$standard.errors
pvals <- (1 - pnorm(abs(zvals), 0, 1)) * 2

# and then display the exponentiated coefficient along with p values afterwards
exp(coef(m1))
pvals

```

Since Labour is the baseline outcome level, all coefficients correspond to a comparison of support for some other party to support for Labour. The raw coefficients represent the log odds of a Labour voting intention vs Conservative. The exponentiated form is the ratio of the probabilities of choosing one outcome category over the other, baseline category (in this case, Labour).

Looking at the coefficients for Conservative, we see that lower education levels are associated with greater support for the Conservatives, relative to Labour (positive coefficients for education levels below BA which translates to values above 1 when they are exponentiated).  There is not much different between the various education levels that are less than a BA, which is consistent with the cross-tabulation results in question 1. What this means, is that the gap between Labour and Conservative voting intention is larger at the lower education levels than it is at the BA level. 

Looking at the coefficients for Liberal Democrat, we see that lower education levels are associated with lower levels of support for the Liberal Democrats, relative to Labour (increasingly negative raw coefficients/values below 1 when exponentiated as education levels go down).  In this case, this means that the lower education levels, the gap between Labour and Lib Dem voters increases, in favour of Labour.

Looking at the coefficients for Other, we see that education levels below BA are associated with higher levels of support for other parties, but there is not a clear trend among the lower education levels - i.e. that the gap between the parties is generally smaller at high education levels, but there is not clear difference across the education levels.   

All of these patterns are consistent with what we see in the cross-tabulation id you look at the 'gaps' between each parties vote intention share at each education level and compare them to the gap at the highest education level.




```{r notes}

ukvote2010$temp[ukvote2010$VoteIntention == "Labour"] <- 0
ukvote2010$temp[ukvote2010$VoteIntention == "Conservative"] <- 1

m2 <- glm(temp ~ Qualifications, data = ukvote2010[ukvote2010$VoteIntention == "Labour"| 
                                                              ukvote2010$VoteIntention =="Conservative",])

summary(m2)
exp(coef(m2))

```



