---
title: "SO5012 Analysising Data in the Real World  \nMultinomial regression  \nSolutions and commentary"
output: 
  pdf_document: 
    toc: yes
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list = ls())

#whereeveryousavestuff <- "a place"
# Note to RP - Just sneakily saving a variable that has the directory path so make the code look better below
```

## Introduction

This seminar sheet is intended as a introduction to multinomial regression and is a combination of code and **inpterpreation** for the worksheet *SO5012_semX_multinomial_regression.docx*. 

As with all the previous weeks, first we need to:
1. Set the working directory
2. Load the packages we'll be using
3. Load the data

Here we'll use the `echo = false` command on the r chunk so that our output is not filled up by this set up code, although the code will be vivible. 

```{r loading, echo=FALSE}

# setwd(whereeveryousavestuff)
# Note to RP - this isnt needed for a project, but you'll need to change if just posting rmd

if (!require(nnet)) install.packages("nnet")
library(nnet)

ukvote2010 <- read.csv("data/ukvote2010.csv")


```


