---
title: "SO5012 Analysising Data in the Real World  \nMultinomial regression  \nSolutions and commentary"
output: 
  pdf_document: 
    toc: yes
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list = ls())

#whereeveryousavestuff <- "a place"
# Note to RP - Just sneakily saving a variable that has the directory path so make the code look better below
```

## Introduction

This seminar sheet is intended as a introduction to multinomial regression and is a combination of code and **interpretation** for the worksheet *SO5012_semX_multinomial_regression.docx*. 

As with all the previous weeks, first we need to:
1. Set the working directory
2. Load the packages we'll be using
3. Load the data

Here we'll use the `results='hide', message=FALSE` command on the r chunk so that our output is not filled up by this set up code, although the code will be visible. 

```{r loading, results='hide', message=FALSE}

# setwd(whereeveryousavestuff)
# Note to RP - this isnt needed for a project, but you'll need to change if just posting rmd

if (!require(nnet)) install.packages("nnet")
library(nnet)

ukvote2010 <- read.csv("data/ukvote2010.csv")

```

To continue with the preparatory work, and **before** we dive into answering the questions, we first need to look at our data and see what it is, perhaps with some data cleaning if needed.

```{r inspection and cleaning}

str(ukvote2010)

names(ukvote2010)
# we c an that "VoteIntention", "Region", "Gender", "Ethnicity", "MaritalStatus", 
# "Housing", "Qualifications" are character variables - lets set them to factors

# and lets be smart and do it in a loop

factorvars <- c("VoteIntention", "Region", "Gender", "Ethnicity", 
                "MaritalStatus", "Housing", "Qualifications")
for (v in factorvars) {
  ukvote2010[[v]] <- as.factor(ukvote2010[[v]])
}


```

